### 2024-07-26 docs/SOSP 2024/rebuttal.txt

Q1: Novelty of scheduling language.

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for this domain and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like tree-walk interleaving, tree-walk unrolling, reduction optimizations etc. These cannot be supported with existing scheduling languages. This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding a few required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work (e.g. Tahoe, RAPIDs) has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

@@

Q2: Comparison with Hummingbird

Hummingbird translates decision tree inference into a series of tensor operations to use PyTorch. While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations. The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction. Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks). This, as discussed in our paper, is not always useful (Section 2). However. it is very difficult to express other SilvanForge optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

@@

The tensor compiler cannot perform the required caching and reduction optimizations either. Since the generated tensor code relies on gathering individual features when required, the indirect accesses to trees and rows complicates the analysis of memory access and reuse. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.

The above discussion brings up a more practical system level problem. Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower level details like shared memory, thread synchronization, etc. Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized. Systems like TorchDynamo and TorchInductor do not provide this level of control to the PyTorch program.

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap.

Rev#B: Are your compiler changes 'portable' across platforms?
The scheduling language, high-level IR and mid-level IR and optimizations on them are fully shared. The same scheduling language is used to specify the structure of inference code on CPUs and GPUs. When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated. The compilation pipelines diverge starting at the lowering of MIR to low-level IR. Even so, we are able to share significant amount of code between the CPU and GPU lowering because of how in-memory representations are abstracted (Section 6).

### 2024-07-26 docs/SOSP 2024/rebuttal.txt

Further, using tensor operations to represent tree traversals obscures the computation's semantics, making it difficult for the tensor compiler to perform these optimizations. Consider the example of walk-interleaving. SilvanForge’s tree-walk interleaving relies on unrolling only the trees whose walks are being interleaved to the same depth. It also reorders trees so that trees with the same depth are processed together. While it maybe possible to extend tensor compilers to perform interleaving, it will be very hard to mimic what SilvanForge does. Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations. Even if this support is added, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.

The tensor compiler cannot perform the required caching and reduction optimizations either. Since the generated tensor code relies on gathering individual features when required, the indirect accesses to trees and rows complicates the analysis of memory access and reuse. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.

The above discussion brings up a more practical system level problem. Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower level details like shared memory, thread synchronization, etc. Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized. Systems like TorchDynamo and TorchInductor do not provide this level of control to the PyTorch program.

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap. We expect that a future comparison of Hummingbird and SilvanForge on GPUs will show similar results.

Rev#D Practical applicability.
There is interest from industry practitioners to adopt SilvanForge based on the results we demonstrate. We gather that there are a wide variety of application specific use-cases where maximizing performance at both low and high batch sizes is important.

### 2024-07-26 docs/SOSP 2024/rebuttal.txt

Further, using tensor operations to represent tree traversals obscures the computation's semantics, making it difficult for the tensor compiler to perform these optimizations. Consider the example of walk-interleaving. SilvanForge’s tree-walk interleaving relies on unrolling only the trees whose walks are being interleaved to the same depth. It also reorders trees so that trees with the same depth are processed together. While it maybe possible to extend tensor compilers to perform interleaving, it will be very hard to mimic what SilvanForge does. Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations. Even if this support is added, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.

The tensor compiler cannot perform the required caching and reduction optimizations either. Since the generated tensor code relies on gathering individual features when required, the indirect accesses to trees and rows complicates the analysis of memory access and reuse. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.

The above discussion brings up a more practical system-level problem. Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower-level details like shared memory, thread synchronization, etc. Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized. Systems like TorchDynamo and TorchInductor do not provide this level of control to the PyTorch program.

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap. We expect that a future comparison of Hummingbird and SilvanForge on GPUs will show similar results.

Rev#D Practical applicability.

Industry practitioners are interested in adopting SilvanForge based on the results we demonstrate. We gather that there are a wide variety of application-specific use-cases where maximizing performance at both low and high batch sizes is important.

Rev#B: Are your compiler changes 'portable' across platforms?

The scheduling language, high-level IR and mid-level IR and optimizations on them are fully shared. The same scheduling language is used to specify the structure of inference code on CPUs and GPUs. When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated. The compilation pipelines diverge starting at the lowering of MIR to low-level IR. Even so, we are able to share a significant amount of code between the CPU and GPU lowering because of how in-memory representations are abstracted (Section 6).

Rev#C: Support for reductions.

MLIR currently supports parallel reductions for value types (scf.reduce) and reducing a dimension of a tensor (linalg.reduce). Neither of these meets the requirements for generating code for decision tree inference. With scf.reduce, every accumulation creates a logical copy and eliminating these copies correctly is a non-trivial problem. With linalg.reduce, the reduction cannot be fused with another loop as the entire tensor of values that are to be reduced need to be written to memory first. The reduction support we implement in SilvanForge is fairly general. Only the process of identifying reduction loops is currently specific to decision tree inference. All lowering of the ops in the reduction dialect are general.

### 2024-07-27 docs/SOSP 2024/rebuttal.txt

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for this domain and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like tree-walk interleaving, tree-walk unrolling, reduction optimizations etc. These cannot be supported with existing scheduling languages. This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains (deep learning vs. image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding a few required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work (e.g. Tahoe, RAPIDs) has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

Q2: Comparison with Hummingbird, PyTorch etc.

Hummingbird translates decision tree inference into a series of tensor operations to use PyTorch. While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations. The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction. Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks). This, as discussed in our paper, is not always useful (Section 2). However, it is very difficult to express other SilvanForge optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.

@@

Further, using tensor operations to represent tree traversals obscures the computation's semantics, making it difficult for the tensor compiler to perform these optimizations. Consider the example of walk-interleaving. SilvanForge’s tree-walk interleaving relies on unrolling only the trees whose walks are being interleaved to the same depth. It also reorders trees so that trees with the same depth are processed together. While it maybe possible to extend tensor compilers to perform interleaving, it will be very hard to mimic what SilvanForge does. Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations. Even if this support is added, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.

The tensor compiler cannot perform the required caching and reduction optimizations either. Since the generated tensor code relies on gathering individual features when required, the indirect accesses to trees and rows complicates the analysis of memory access and reuse. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.

The above discussion brings up a more practical system-level problem. Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower-level details like shared memory, thread synchronization, etc. Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized. Systems like TorchDynamo and the recently introduced TorchInductor (as per the ASPLOS 2024 paper and other docs) do not provide this level of control to the PyTorch program.

### 2024-07-27 docs/SOSP 2024/rebuttal.txt

We thank the reviewers for their detailed and insightful reviews.

_Q1: Novelty of scheduling language._

@@

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap. We expect that a future comparison of Hummingbird and SilvanForge on GPUs will show similar results.

_Rev#D: Practical applicability._

Industry practitioners are interested in adopting SilvanForge based on the results we demonstrate. We have an ongoing discussion with relevant stakeholders. We gather that there are a wide variety of application-specific use-cases where maximizing performance at both low and high batch sizes is important.

_Rev#A: Benchmarks, absolute times._

We used the benchmark set from previous work and included a large number of synthetic models with variations in depth, number of trees, and number of features to provide additional coverage. Our results are consistent across all these models, indicating robustness. Regarding the execution times, the reported numbers represent the amortized execution time per row averaged over the batch. On a GPU, multiple operations are performed in parallel, making it challenging to directly correlate the reported times with single operations.

_Rev#B: Portability._

### 2024-07-27 docs/SOSP 2024/rebuttal.txt

We thank the reviewers for their insightful reviews.

_Q1: Novelty of scheduling language._

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for this domain and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like tree-walk interleaving, tree-walk unrolling, reduction optimizations etc. These cannot be supported with existing scheduling languages. This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains, such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding a few required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

@@

_Q2: Comparison with Hummingbird, PyTorch etc._

Hummingbird translates decision tree inference into a series of tensor operations to use PyTorch. While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations. The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction. Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks), which is not always useful (Section 2). However, it is very difficult to express other SilvanForge optimizations including tree walk interleaving, caching and reduction optimizations, using tensor operations.

@@

Further, using tensor operations to represent tree traversals obscures the computation's semantics, making it difficult for the tensor compiler to perform these optimizations. Consider the example of walk-interleaving. SilvanForge’s tree-walk interleaving relies on unrolling only the trees whose walks are being interleaved to the same depth. It also reorders trees so that trees with the same depth are processed together. While it maybe possible to extend tensor compilers to perform interleaving, it will be very hard to mimic what SilvanForge does. Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations. Even if this support is added, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.

The tensor compiler cannot perform the required caching and reduction optimizations either. Since the generated tensor code relies on gathering individual features when required, the indirect accesses to trees and rows complicates the analysis of memory access and reuse. This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does. Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walks and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.

The above discussion brings up a more practical system-level problem. Expressing inference using PyTorch means that the decision tree compiler has no control over lower-level details like shared memory, thread synchronization, etc. Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized. Systems like TorchDynamo and the recently introduced TorchInductor (as per the ASPLOS 2024 paper and other docs) do not provide this level of control to the PyTorch program.

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap. We expect a future comparison of Hummingbird and SilvanForge on GPUs to show similar results.

@@

_Rev#D: Practical applicability._

Industry practitioners are interested in adopting SilvanForge based on the results we demonstrate. We have an ongoing discussion with relevant stakeholders. We gather that there are a wide variety of application-specific use-cases where maximizing performance at both low and high batch sizes is important.

_Rev#A: Benchmarks._

We used the benchmark set from previous work and included a large number of synthetic models with variations in depth, number of trees, and number of features to provide additional coverage. Our results are consistent across all these models, indicating robustness.

_Rev#A: Absolute times._

The reported numbers are the mean execution time per row averaged over the batch. Since several rows are processed in parallel on the GPU, it is challenging to directly correlate the reported times with single operations.

@@

_Rev#B: Portability._

The scheduling language, HIR and MIR and optimizations on them are fully shared. The same scheduling language is used to specify the structure of inference code on CPUs and GPUs. When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated. The compilation pipelines diverge starting at the lowering of MIR to LIR. Even so, we are able to share a significant amount of code between the CPU and GPU lowering because of how in-memory representations are abstracted (Section 6).

@@

_Rev#C: Support for reductions._

MLIR currently supports parallel reductions for value types (scf.reduce) and reducing dimensions of tensors (linalg.reduce). Neither of these meets the requirements for generating code for decision tree inference. With scf.reduce, every accumulation creates a logical copy and eliminating these copies correctly is a non-trivial problem. With linalg.reduce, the reduction cannot be fused with another loop as the entire tensor of values that are to be reduced need to be written to memory first. The reduction support we implement in SilvanForge is fairly general. Only the process of identifying reduction loops is currently specific to decision tree inference. All lowering of the ops in the reduction dialect are general.

### 2024-07-27 docs/SOSP 2024/rebuttal.txt

Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees. One of the paper's key contributions lies in recognizing the existence of a scheduling space for this domain and demonstrating the significant performance variations across different schedules. The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like tree-walk interleaving, tree-walk unrolling, reduction optimizations etc. These cannot be supported with existing scheduling languages. This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading. Despite the fact that TVM and Halide operate on similar domains, such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.

We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding a few required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference. Prior work has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.

_Q2: Comparison with Hummingbird, PyTorch etc._

Hummingbird translates decision tree inference into a series of tensor operations to use PyTorch. While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations. In the latest Hummingbird version, the generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction. Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks), which is not always useful (Section 2). However, it is very difficult to express other SilvanForge optimizations including tree walk interleaving, caching and reduction optimizations, using tensor operations.

@@

The above discussion brings up a more practical system-level problem. Expressing inference using PyTorch means that the decision tree compiler has no control over lower-level details like shared memory, thread synchronization, etc. Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized. Systems like TorchDynamo and the recently introduced TorchInductor (as per the ASPLOS 2024 paper and other docs) do not provide this level of control to the PyTorch program.

These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap. We expect a future comparison of Hummingbird and SilvanForge on GPUs to show similar results.

_Rev#D: Practical applicability._

We are currently engaging with industry practitioners and library code owners who are keen to adopt SilvanForge, inspired by the promising results we’ve demonstrated. Their requirements include maximizing performance across a range of batch sizes, as their application use-cases vary.
